---
title: "Bm_final_project"
author: "Yi Xiao"
date: "10/12/2018"
output: html_document
---
```{r}
library(tidyverse)
library(kableExtra)
library(Hmisc)
```

1. load the dataset
```{r}
cancer_data <- read_csv("./data/Cancer_Registry.csv") %>%
        janitor::clean_names()
```
The Cancer Registry Dataset cantains information on cancer mortality rate across `r n_distinct(cancer_data$geography)` US counties. There are `r ncol(cancer_data)` variables which provide information on death rate and related demographical characteristics in each county. The outcome variable in this datset is `target_death rate` (continuous). 

2. explore the data

```{r}
# 1. deal with missing data

num_var <- cancer_data %>%
        select(-c(binned_inc, geography))
# summary table for all numberical variables:
num_var %>%
        skimr::skim_to_wide() %>% knitr::kable() %>% kable_styling(full_width = F, position = "left")

cancer_data %>%
   select(geography,binned_inc) %>%
describe()

# presence of large amount of missing data (over 20%)  in two variabls: pct_some_col18_24 and pct_private_coverage_alone
# remove those two variables 

```
 
 
```{r}
# 2. check normality
ggplot(gather(num_var), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
```
 
heavyly-left-skewed: avg_ann_count, avg_deaths_per_year, med_income, pop_est2015, avg_ann_count, median_age, pct_asian, pct_black, pct_other_race, pop_est2015, study_per_cap
right-skewed: pct_white
outcome variable has a roughly normal disribution.

3. tidy
```{r}
cancer_tidy_data = cancer_data %>% 
        separate(geography, into = c("county", "state"), sep = ", ") %>%
        mutate(
                binned_inc_lowerb = str_split_fixed(binned_inc, ", ", 2)[ ,1] %>% parse_number(), 
                binned_inc_upperb = str_split_fixed(binned_inc, ", ", 2)[ ,2] %>% parse_number(), 
                binned_inc_mean = (binned_inc_lowerb + binned_inc_upperb)/2
                ) %>%# calculate point estimate for variable binned_inc  
         mutate(region = case_when(
            state %in% c("Montana", "Arizona","Idaho","Colorado", "New     Mexico","Utah","Nevada","Wyoming","Alaska","California",
            "Hawaii","Oregon","Washington") ~ "West",
            state %in% c("Illinois","Indiana", "Michigan","Iowa","Ohio", "Kansas","Minnesota","Missouri","Wisconsin", "Nebraska","North Dakota","South Dakota") ~ "Midwest",
            state %in% c("District of Columbia", "Delaware","Georgia",
            "Maryland", "North Carolina","Florida","South Carolina",
            "West Virginia","Kentucky", "Alabama","Virginia","Mississippi", "Arkansas","Tennessee","Louisiana","Oklahoma","Texas") ~ "South",
             state %in%  c("Connecticut","Maine","Massachusetts","New Hampshire",
             "Rhode Island","Vermont","New Jersey","New York",
             "Pennsylvania") ~ "Northeast"
                )) 
```

