---
title: "Bm_final_project"
author: "Yi Xiao"
date: "10/12/2018"
output: html_document
---
```{r}
library(tidyverse)
library(kableExtra)
library(Hmisc)
```

1. load the dataset
```{r}
cancer_data <- read_csv("./data/Cancer_Registry.csv") %>%
        janitor::clean_names()

head(cancer_data)
str(cancer_data)

# number of recordings
nrow(cancer_data)

# number of variables
ncol(cancer_data)

```
The Cancer Registry Dataset cantains information on cancer mortality rate across `r n_distinct(cancer_data$geography)` US counties. There are `r ncol(cancer_data)` variables which provide information on death rate and related demographical characteristics in each county. The outcome variable in this datset is `target_death rate` (continuous). 

2. explore the data

```{r}
# 1. deal with missing data

num_var <- cancer_data %>%
        select(-c(binned_inc, geography))
# summary table for all numberical variables:
num_var %>%
        skimr::skim_to_wide() %>% knitr::kable() %>% kable_styling(full_width = F, position = "left")

cancer_data %>%
   select(geography,binned_inc) %>%
describe()

# presence of large amount of missing data (over 20%)  in two variabls: pct_some_col18_24 and pct_private_coverage_alone
# remove those two variables 

```
 
 
```{r}
# 2. check normality
ggplot(gather(num_var), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
```
 
heavyly-left-skewed: avg_ann_count, avg_deaths_per_year, med_income, pop_est2015, avg_ann_count, median_age, pct_asian, pct_black, pct_other_race, pop_est2015, study_per_cap
right-skewed: pct_white
outcome variable has a roughly normal disribution.

