---
title: "sammi_methods_final_project"
author: "Samantha Brown"
date: "12/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(Hmisc)
```

**Load Data**
```{r, message = FALSE}
cancer_data <- read_csv("./data/Cancer_Registry.csv") %>%
        janitor::clean_names()
```

### Data Exploration

1. `r round(cancer_data %>% mutate(na_data = is.na(pct_some_col18_24)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_some_col18_24` is missing. Therefore, we will remove it from the dataset.

2. `r round(cancer_data %>% mutate(na_data = is.na(pct_private_coverage_alone)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_private_coverage_alone` is missing. Therefore, we will remove it from the dataset.

3. `med_income` and `poverty_percent` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$med_income, cancer_data$poverty_percent)*100, digits = 2)`%. To account for this, we will remove `poverty_percent` from the dataset.

4. `median_age_female` and `median_age_male` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$median_age_female, cancer_data$median_age_male)*100, digits = 2)`%. Thus, we will only consider `median_age` for our data analysis.

5. `avg_ann_count` and `avg_deaths_per_year` have a high correlation coefficient of `r round(cor(cancer_data$avg_ann_count, cancer_data$avg_deaths_per_year)*100, digits = 2)`%. Both of these variables are also highly correlated with `pop_est2015`. We will remove `avg_ann_count` and `avg_deaths_per_year`.

```{r}
cor(cancer_data$pct_private_coverage, cancer_data$pct_public_coverage)
cor(cancer_data$pct_emp_priv_coverage, cancer_data$pct_private_coverage)
cor(cancer_data$pct_public_coverage_alone, cancer_data$pct_public_coverage)
```


```{r, include = FALSE}
cor(cancer_data$median_age, cancer_data$median_age_female)
cor(cancer_data$median_age, cancer_data$median_age_male)
cor(cancer_data$median_age_female, cancer_data$median_age_male)
```

```{r}
cancer_data = cancer_data %>% 
  select(-pct_some_col18_24, -pct_private_coverage_alone, -poverty_percent, -median_age_female, -median_age_male, -avg_ann_count, -avg_deaths_per_year, -pct_public_coverage, -pct_hs25_over, -pct_employed16_over) %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>%
        mutate(
                binned_inc_lowerb = str_split_fixed(binned_inc, ", ", 2)[ ,1] %>% parse_number(), 
                binned_inc_upperb = str_split_fixed(binned_inc, ", ", 2)[ ,2] %>% parse_number(), 
                binned_inc_mean = (binned_inc_lowerb + binned_inc_upperb)/2)
```

Start with full model
```{r}
mult.fit <- lm(target_death_rate ~ ., data = cancer_data)
```

```{r}

```


