---
title: "sammi_methods_final_project"
author: "Samantha Brown"
date: "12/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(Hmisc)
```

**Load Data**
```{r, message = FALSE}
cancer_data <- read_csv("./data/Cancer_Registry.csv") %>%
        janitor::clean_names()
```

### Data Exploration

1. `r round(cancer_data %>% mutate(na_data = is.na(pct_some_col18_24)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_some_col18_24` is missing. Therefore, we will remove it from the dataset.

2. `r round(cancer_data %>% mutate(na_data = is.na(pct_private_coverage_alone)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_private_coverage_alone` is missing. Therefore, we will remove it from the dataset.

3. `med_income` and `poverty_percent` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$med_income, cancer_data$poverty_percent)*100, digits = 2)`%. To account for this, we will remove `poverty_percent` from the dataset.

4. `median_age_female` and `median_age_male` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$median_age_female, cancer_data$median_age_male)*100, digits = 2)`%. Thus, we will only consider `median_age` for our data analysis.

5. `avg_ann_count` and `avg_deaths_per_year` have a high correlation coefficient of `r round(cor(cancer_data$avg_ann_count, cancer_data$avg_deaths_per_year)*100, digits = 2)`%. Both of these variables are also highly correlated with `pop_est2015`. We will remove `avg_ann_count` and `avg_deaths_per_year`.

```{r}
cor(cancer_data$pct_private_coverage, cancer_data$pct_public_coverage)
cor(cancer_data$pct_emp_priv_coverage, cancer_data$pct_private_coverage)
cor(cancer_data$pct_public_coverage_alone, cancer_data$pct_public_coverage)
```


```{r, include = FALSE}
cor(cancer_data$median_age, cancer_data$median_age_female)
cor(cancer_data$median_age, cancer_data$median_age_male)
cor(cancer_data$median_age_female, cancer_data$median_age_male)
```

```{r}
tidy_data = cancer_data %>% 
  select(-pct_some_col18_24, -pct_private_coverage_alone, -poverty_percent, -median_age_female, -median_age_male, -avg_ann_count, -avg_deaths_per_year, -pct_public_coverage, -pct_hs25_over, -pct_employed16_over, -pct_married_households, -pct_emp_priv_coverage, -binned_inc, -pct_hs18_24, -pct_no_hs18_24, -pct_bach_deg18_24) %>% 
  separate(geography, into = c("county", "state"), sep = ", ") %>% 
   mutate(region = case_when(  
            state %in% c("Montana", "Arizona","Idaho","Colorado", "New Mexico","Utah","Nevada","Wyoming","Alaska","California",
            "Hawaii","Oregon","Washington") ~ "West",
            state %in% c("Illinois","Indiana", "Michigan","Iowa","Ohio", "Kansas","Minnesota","Missouri","Wisconsin", "Nebraska","North Dakota","South Dakota") ~ "Midwest",
            state %in% c("District of Columbia", "Delaware","Georgia",
            "Maryland", "North Carolina","Florida","South Carolina",
            "West Virginia","Kentucky", "Alabama","Virginia","Mississippi", "Arkansas","Tennessee","Louisiana","Oklahoma","Texas") ~ "South",
             state %in%  c("Connecticut","Maine","Massachusetts","New Hampshire",
             "Rhode Island","Vermont","New Jersey","New York",
             "Pennsylvania") ~ "Northeast" ), region = as.factor(region) %>% fct_relevel(., "West", "Midwest", "Northeast", "South"))

tidy_data = tidy_data %>%
  select(-county, -state) %>% 
  mutate(pct_non_black = pct_asian + pct_white + pct_other_race) %>% 
  select(-pct_asian, -pct_white, -pct_other_race)
        
```

Start with full model
```{r}
mult.fit <- lm(target_death_rate ~ ., data = tidy_data)
```

```{r}
library(broom)

## Forward elimination
## Run regression on all individual variables 

fit1 = lm(target_death_rate ~ incidence_rate, data = tidy_data)
tidy(fit1)

fit2 = lm(target_death_rate ~ med_income, data = tidy_data)
tidy(fit2)

fit3 = lm(target_death_rate ~ pop_est2015, data = tidy_data)
tidy(fit3)

fit4 = lm(target_death_rate ~ study_per_cap, data = tidy_data)
tidy(fit4)

fit5 = lm(target_death_rate ~ median_age, data = tidy_data)
tidy(fit5)

fit6 <- lm(target_death_rate ~ avg_household_size, data = tidy_data)
tidy(fit6)

fit7 <- lm(target_death_rate ~ percent_married, data = tidy_data)
tidy(fit7)

fit8 = lm(target_death_rate ~ pct_bach_deg25_over, data = tidy_data)
tidy(fit8)

fit9 = lm(target_death_rate ~ pct_unemployed16_over, data = tidy_data)
tidy(fit9)

fit10 = lm(target_death_rate ~ pct_private_coverage, data = tidy_data)
tidy(fit10)

fit11 = lm(target_death_rate ~ pct_public_coverage_alone, data = tidy_data)
tidy(fit11)

fit12 = lm(target_death_rate ~ pct_black, data = tidy_data)
tidy(fit12)

fit13 <- lm(target_death_rate ~ pct_non_black, data = tidy_data)
tidy(fit13)

fit14 <- lm(target_death_rate ~ birth_rate, data = tidy_data)
tidy(fit14)

fit15 <- lm(target_death_rate ~ region, data = tidy_data)
tidy(fit15)

## Begin forward elimination -- Step 1

forward1 = lm(target_death_rate ~ pct_bach_deg25_over, data = tidy_data)
tidy(forward1)

fit1 = update(forward1, . ~ . +incidence_rate)
tidy(fit1)

fit2 = update(forward1, . ~ . +med_income)
tidy(fit2)

fit3 = update(forward1, . ~ . +pop_est2015)
tidy(fit3)

fit4 = update(forward1, . ~ . +study_per_cap)
tidy(fit4)

fit5 = update(forward1, . ~ . +median_age)
tidy(fit5)

fit6 = update(forward1, . ~ . +avg_household_size)
tidy(fit6)

fit7 = update(forward1, . ~ . +percent_married)
tidy(fit7)

fit8 = update(forward1, . ~ . +pct_unemployed16_over)
tidy(fit8)

fit9 = update(forward1, . ~ . +pct_private_coverage)
tidy(fit9)

fit10 = update(forward1, . ~ . +pct_public_coverage_alone)
tidy(fit10)

fit11 = update(forward1, . ~ . +pct_black)
tidy(fit11)

fit12 = update(forward1, . ~ . +pct_non_black)
tidy(fit12)

fit13 = update(forward1, . ~ . +birth_rate)
tidy(fit13)

fit14 = update(forward1, . ~ . +region)
tidy(fit14)

## Step 2

forward2 <- update(forward1, . ~ . +incidence_rate)
tidy(forward2)

fit1 = update(forward2, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward2, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward2, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward2, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward2, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward2, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward2, . ~ . +pct_unemployed16_over)
tidy(fit7)

fit8 = update(forward2, . ~ . +pct_private_coverage)
tidy(fit8)

fit9 = update(forward2, . ~ . +pct_public_coverage_alone)
tidy(fit9)

fit10 = update(forward2, . ~ . +pct_black)
tidy(fit10)

fit11 = update(forward2, . ~ . +pct_non_black)
tidy(fit11)

fit12 = update(forward2, . ~ . +birth_rate)
tidy(fit12)

fit13 = update(forward2, . ~ . +region)
tidy(fit13)

## Step 3
forward3 = update(forward2, . ~ . + pct_private_coverage)
tidy(forward3)

fit1 = update(forward3, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward3, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward3, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward3, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward3, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward3, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward3, . ~ . +pct_unemployed16_over)
tidy(fit7)

fit8 = update(forward3, . ~ . +pct_public_coverage_alone)
tidy(fit8)

fit9 = update(forward3, . ~ . +pct_black)
tidy(fit9)

fit10 = update(forward3, . ~ . +pct_non_black)
tidy(fit10)

fit11 = update(forward3, . ~ . +birth_rate)
tidy(fit11)

fit12 = update(forward3, . ~ . +region)
tidy(fit12)

## Step 4
forward4 = update(forward3, . ~ . + region)
tidy(forward4)

fit1 = update(forward4, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward4, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward4, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward4, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward4, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward4, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward4, . ~ . +pct_unemployed16_over)
tidy(fit7)

fit8 = update(forward4, . ~ . +pct_public_coverage_alone)
tidy(fit8)

fit9 = update(forward4, . ~ . +pct_black)
tidy(fit9)

fit10 = update(forward4, . ~ . +pct_non_black)
tidy(fit10)

fit11 = update(forward4, . ~ . +birth_rate)
tidy(fit11)

## Step 5
forward5 = update(forward4, . ~ . + pct_unemployed16_over)
tidy(forward5)

fit1 = update(forward5, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward5, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward5, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward5, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward5, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward5, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward5, . ~ . +pct_public_coverage_alone)
tidy(fit7)

fit8 = update(forward5, . ~ . +pct_black)
tidy(fit8)

fit9 = update(forward5, . ~ . +pct_non_black)
tidy(fit9)

fit10 = update(forward5, . ~ . +birth_rate)
tidy(fit10)

## Step 6

forward6 = update(forward5, . ~ . + pct_public_coverage_alone)
tidy(forward6)

fit1 = update(forward6, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward6, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward6, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward6, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward6, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward6, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward6, . ~ . +pct_black)
tidy(fit7)

fit8 = update(forward6, . ~ . +pct_non_black)
tidy(fit8)

fit9 = update(forward6, . ~ . +birth_rate)
tidy(fit9)

## Step 7

forward7 = update(forward6, . ~ . + birth_rate)
tidy(forward7)

fit1 = update(forward7, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward7, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward7, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward7, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward7, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward7, . ~ . +percent_married)
tidy(fit6)

fit7 = update(forward7, . ~ . +pct_black)
tidy(fit7)

fit8 = update(forward7, . ~ . +pct_non_black)
tidy(fit8)

## Step 8

forward8 = update(forward7, . ~ . + percent_married)
tidy(forward8)

fit1 = update(forward8, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward8, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward8, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward8, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward8, . ~ . +avg_household_size)
tidy(fit5)

fit6 = update(forward8, . ~ . +pct_black)
tidy(fit6)

fit7 = update(forward8, . ~ . +pct_non_black)
tidy(fit7)

## Step 9

forward9 = update(forward8, . ~ . + pct_black)
tidy(forward9)

fit1 = update(forward9, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward9, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward9, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward9, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward9, . ~ . +avg_household_size)
tidy(fit5)

fit8 = update(forward9, . ~ . +pct_non_black)
tidy(fit8)


## Step 10

forward10 = update(forward9, . ~ . + pct_non_black)
tidy(forward10)

fit1 = update(forward10, . ~ . +med_income)
tidy(fit1)

fit2 = update(forward10, . ~ . +pop_est2015)
tidy(fit2)

fit3 = update(forward10, . ~ . +study_per_cap)
tidy(fit3)

fit4 = update(forward10, . ~ . +median_age)
tidy(fit4)

fit5 = update(forward10, . ~ . +avg_household_size)
tidy(fit5)

## Predicted model from forward elimination: 
mult_forward_fit = lm(target_death_rate ~ pct_bach_deg25_over + incidence_rate + pct_private_coverage + region + pct_unemployed16_over + pct_public_coverage_alone + birth_rate + percent_married + pct_black + pct_non_black, data = tidy_data)

tidy(mult_forward_fit)
```


