---
title: "sammi_methods_final_project"
author: "Samantha Brown"
date: "12/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(Hmisc)
```

Load Data
```{r, message = FALSE}
cancer_data <- read_csv("./data/Cancer_Registry.csv") %>%
        janitor::clean_names()
```

### Data Exploration

* `r round(cancer_data %>% mutate(na_data = is.na(pct_some_col18_24)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_some_col18_24` is missing. Therefore, we will remove it from the dataset.

* `r round(cancer_data %>% mutate(na_data = is.na(pct_private_coverage_alone)) %>% count(na_data) %>% filter(na_data == TRUE) %>% select(n)/nrow(cancer_data)*100, digits = 1)`% of the variable `pct_private_coverage_alone` is missing. Therefore, we will remove it from the dataset.

* `med_income` and `poverty_percent` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$med_income, cancer_data$poverty_percent)*100, digits = 2)`%. To account for this, we will remove `poverty_percent` from the dataset.

* `median_age_female` and `median_age_male` are highly correlated with a correlation coefficient of `r round(cor(cancer_data$median_age_female, cancer_data$median_age_male)*100, digits = 2)`%. Thus, we will only consider `median_age` for our data analysis.


```{r, include = FALSE}
cor(cancer_data$median_age, cancer_data$median_age_female)
cor(cancer_data$median_age, cancer_data$median_age_male)
cor(cancer_data$median_age_female, cancer_data$median_age_male)
```


```{r}
cancer_data = cancer_data %>% 
  select(-pct_some_col18_24, -pct_private_coverage_alone, -poverty_percent, -median_age_female, -median_age_male)
```

Start with full model
```{r}
##mult.fit <- lm(target_death_rate ~ ., data = cancer_data)
##summary(mult.fit)
```



